---
title: "Final Project EDAV"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# EDAV Final Project
## Data Loading and Cleaning
```{r message=FALSE}
library(tidyverse)

# Load data
student_mat <- read_csv("/Users/naoto/Downloads/student-alcohol-consumption/student-mat.csv")
student_por <- read_csv("/Users/naoto/Downloads/student-alcohol-consumption/student-por.csv")

# Merge student in math course and portuguese language course
student <- student_mat %>%
  full_join(student_por,
            by = c("school","sex","age","address","famsize",
                   "Pstatus","Medu","Fedu","Mjob","Fjob",
                   "reason","nursery","internet")) %>%
  mutate(G3=if_else(is.na(`G3.x`), `G3.y`, `G3.x`)) %>%
  mutate(guardian=if_else(is.na(`guardian.x`), `guardian.y`, `guardian.x`)) %>%
  mutate(absences=if_else(is.na(`absences.x`), `absences.y`, `absences.x`)) %>%
  mutate(failures=if_else(is.na(`failures.x`), `failures.y`, `failures.x`)) %>%
  mutate(activities=if_else(is.na(`activities.x`), `activities.y`, `activities.x`)) %>%
  #mutate(course=if_else(is.na(`G3.x`), "por", "mat")) %>%
  mutate(FinalGrade = if_else(G3>=16, 'excellent',
                            if_else(G3>=14&G3<=15, 'good',
                                    if_else(G3>=12&G3<=13, 'satisfactory',
                                            if_else(G3>=10&G3<=11, 'sufficient',
                                                    'fail'))))) %>%
  mutate(FinalGrade = factor(FinalGrade,
                             c('excellent',
                               'good',
                               'satisfactory',
                               'sufficient',
                               'fail'))) %>%
  mutate(Medu=recode(Medu,
                      `0`='none',
                      `1`='primary education',
                      `2`='5th to 9th grade',
                      `3`='secondary education',
                      `4`='higher education')) %>%
  mutate(Fedu=recode(Fedu,
                      `0`='none',
                      `1`='primary education',
                      `2`='5th to 9th grade',
                      `3`='secondary education',
                      `4`='higher education')) %>%
  mutate(Medu=factor(Medu,
                     c('none',
                       'primary education',
                       '5th to 9th grade',
                       'secondary education',
                       'higher education'))) %>%
  mutate(Fedu=factor(Fedu,
                     c('none',
                       'primary education',
                       '5th to 9th grade',
                       'secondary education',
                       'higher education')))
```

## Association between Father's job and Students' Score
If father's job is teacher, students' grades tend to be higher.
```{r message=FALSE}
library(vcd)
vcd::mosaic(FinalGrade~Fjob, student, rot_labels=c(30,0,0,0), direction=c("v","h"))
```

```{r message=FALSE}
ggplot(student, aes(x = reorder(`Fjob`, G3, median),
                    y = G3)) + 
  geom_boxplot(varwidth = TRUE) +
  coord_flip() +
  theme_grey(14)
```

## Association between Mother's job and Students' Score
If Mother's job is health, students' grades tend to be higher.
```{r message=FALSE}
library(vcd)
vcd::mosaic(FinalGrade~Mjob, student, rot_labels=c(30,0,0,0), direction=c("v","h"))
```

```{r message=FALSE}
ggplot(student, aes(x = reorder(`Mjob`, G3, median),
                    y = G3)) + 
  geom_boxplot(varwidth = TRUE) +
  coord_flip() +
  theme_grey(14)
```

## Association between Parents' job and Students' Score
What if both parents are at home.
```{r message=FALSE}
student <- student %>%
  mutate(Mjob2=if_else(Mjob=='at_home', 'at_home','working')) %>%
  mutate(Fjob2=if_else(Fjob=='at_home', 'at_home','working'))

library(vcd)
vcd::mosaic(FinalGrade~Fjob2+Mjob2, student, rot_labels=c(30,0,30,0), direction=c("v","v","h"))
```

## Association between Father's education and Students' Score
If father took higher education, students' grades tend to be higher.
```{r message=FALSE}
library(vcd)
vcd::mosaic(FinalGrade~Fedu, student, rot_labels=c(30,0,0,0), direction=c("v","h"))
```

```{r message=FALSE}
ggplot(student, aes(x = reorder(`Fedu`, G3, median),
                    y = G3)) + 
  geom_boxplot(varwidth = TRUE) +
  coord_flip() +
  theme_grey(14)
```

## Association between Mother's education and Students' Score
If Mother took higher education, students' grades tend to be higher.
```{r message=FALSE}
library(vcd)
vcd::mosaic(FinalGrade~Medu, student, rot_labels=c(30,0,0,0), direction=c("v","h"))
```

```{r message=FALSE}
ggplot(student, aes(x = reorder(`Medu`, G3, median),
                    y = G3)) + 
  geom_boxplot(varwidth = TRUE) +
  coord_flip() +
  theme_grey(14)
```

## Association between Parents' education and Students' Score
Wen both parents did not take higher eductation, students are less likely to earn excellent and good grades.
```{r message=FALSE}
student <- student %>%
  mutate(Medu2=if_else(Medu=='higher education', 'higher education','other')) %>%
  mutate(Fedu2=if_else(Fedu=='higher education', 'higher education','other'))

library(vcd)
vcd::mosaic(FinalGrade~Fedu2+Medu2, student, rot_labels=c(30,0,30,0), direction=c("v","v","h"))
```


## Association between Absences and Students' Score
### Obtain quantiles for absences
```{r message=FALSE}
# absences quantiles
boxplot(student$absences, horizontal = TRUE, las = 1)
fivenumnames <- c("min", "lower-hinge", "median", "upper-hinge", "max")
D <- student$absences
fivenum(D) %>% set_names(fivenumnames)
text(fivenum(D)[c(1,3,5)], 1.25, round(fivenum(D)[c(1,3,5)],1), col = "red")
text(fivenum(D)[c(2,4)], .75, round(fivenum(D),1)[c(2,4)], col = "red")

# Bucketing absences to 4 qunatiles
student <- student %>%
  mutate(AbsenceFrequency = if_else(absences>=7, 'Most frequent',
                                    if_else(absences>2&absences<7, 'Second frequent',
                                            if_else(absences>0&absences<=2, 'Average',
                                                    if_else(absences==0, 'No Absences',
                                                            'Other'))))) %>%
  mutate(AbsenceFrequency = factor(AbsenceFrequency,
                                   c('Most frequent','Second frequent','Average','No Absences','Other')))

library(vcd)
vcd::mosaic(FinalGrade~AbsenceFrequency, student, rot_labels=c(30,0,0,0), direction=c("v","h"))
```


